---
title: "Sequencing libraries on the NextSeq 1000"
date: |
  As of `r Sys.Date()`
output: 
  bookdown::pdf_document2:
    latex_engine: xelatex
    number_sections: no
    toc_depth: 3
  html_document: default
params: 
  libraries: 4
geometry:
  top=10mm,
  bottom=14mm,
  left=11mm,
  right=11mm
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
```


```{r}
library(tidyverse)
```
\newpage

# Buffers / reagents 

## Qubit 1x dsDNA HS kit

Fisher Q33230

## Qubit assay tubes

Fisher Q32856

## PhiX control library

Illumina FC-110-3001

## NextSeq 1000 reagent kit

See options [here](https://www.illumina.com/products/by-type/sequencing-kits/cluster-gen-sequencing-reagents/nextseq-1000-2000-reagents.html)

# Steps
## Day One

### Thaw the cartridge

1. Remove box containing the cartridge from -20°C.\    

2. Remove the cartridge from the box. **Do not remove cartridge from silver foil bag**\    

3. Position cartridge at room temp so that label faces up and air can circulate on sides and top.\    

4. Thaw at room temperature for 6 hours.\    

5. Transfer to 4°C and thaw for at least 12 hours. **Do not exceed 72 hours**\    

### Set up run on Base Space

6. Navigate to this [page](https://basespace.illumina.com/)

7. Click on **RUNS** within the blue section at the top of the page.\    

8. Click "New run" and select "INSTRUMENT RUN SETUP".\    

9. Fill out the information related to your run.\    

    For paired reads, the number of cycles per read should be $\leq$ (total cycles / 2) + 1.\    
    For example, a P2-300 flow cell can accomodate 150 + 1 cycles per read.\    
    The NEBNext libraries for Illumina resemble TruSeq libraries and can be trimmed like TruSeq:\    
      Adaptor Read1   AGATCGGAAGAGCACACGTCTGAACTCCAGTCA\    
      Adaptor Read2   AGATCGGAAGAGCGTCGTGTAGGGAAAGAGTGT   





## Day Two
### Qubit

*Only use the tubes specific to the Qubit instrument*\    

10. Dilute the high and low standards with 10 $\mu$l of standard : 190 $\mu$l of working buffer.\    

11. Dilute each of your libraries with 1 $\mu$l of library : 199 $\mu$l of working buffer.\    

12. Vortex standards and libraries and then read on the Qubit.\    


```{r qubit}
qubit <- rep("   ", params$libraries) %>% as_tibble_col("library") %>% mutate(
  `ng/$\\mu$l` = .$library
)

knitr::kable(
  qubit, 
  format = "latex", 
  escape = F,
  caption = "Library concentrations as measured by the Qubit",
  align = "c"
) %>% kableExtra::kable_styling(
  latex_options = c("HOLD")
) %>% kableExtra::column_spec(
  c(1:2),
  width = unit(25,"mm")
)
```


### Convert concentrations to molarity

13. Use the following equation to go from ng/$\mu$l to nM.:\    $$[ng/\mu l] *\ 1/660\ g/mol * 1/average\ library \ size\ (bp)\ * 1E6\ = [nM] $$\    

```{r convert}
convert <- qubit %>% mutate(
  nM = .$library
)

knitr::kable(
  convert, 
  format = "latex", 
  escape = F,
  caption = "Library concentrations in nM",
  align = "c"
) %>% kableExtra::kable_styling(
  latex_options = c("HOLD")
) %>% kableExtra::column_spec(
  c(1:3),
  width = unit(25,"mm")
)
```


### Dilute libraries to 2 nM

14. Dilute your libraries to 2 nM using nuclease-free H~2~O


```{r working}
working <- qubit %>% select(
  library
) %>% mutate(
  `Stock (nM)` = .$library
) %>% mutate(
  `Working (nM)` = .$library
)

knitr::kable(
  working, 
  format = "latex", 
  escape = F,
  caption = "Diluting libraries to [Working]",
  align = "c"
) %>% kableExtra::kable_styling(
  latex_options = c("HOLD")
) %>% kableExtra::column_spec(
  c(1:3),
  width = unit(35,"mm")
)
```



### Determine the target concentration for the pool of libraries to be loaded onto the flow cell.\    

```{r}
targets <- c(
  "Ampliseq for Illumina Library PLUS",
  "Illumina DNA Prep",
  "Illumina DNA Prep with Enrichment",
  "Illumina Stranded Total RNA with Ribo-Zero Plus",
  "Illumina Stranded mRNA Prep",
  "Illumina DNA PCR-Free",
  "TruSeq DNA Nano 350a",
  "TruSeq DNA Nano 550",
  "TruSeq Stranded mRNA",
  "100% PhiX"
) %>% as_tibble_col("Library Type") %>% mutate(
  `Loading concentration (pM)` = c(
    750,
    750,
    1000,
    750,
    750,
    1000,
    1200,
    1500,
    1000,
    650
  )
)


knitr::kable(
  targets, 
  caption = "Illumina's suggested loading concentrations for pools containing variously prepared libraries.",
  align = "c",
  format = "latex", 
  booktabs = T
) %>% kableExtra::kable_styling(
  latex_options = c("HOLD", "striped")
) %>% kableExtra::add_footnote(
  label = "The NEBNext libraries for Illumina are analagous to TruSeq libraries."
)
```

15. For this sequencing experiment, we will aim for the following loading concentration **of the pool**: _______\    

    *Pooling concentration is defined by the sum of the molarity of each library within the pool*


### Dilute libraries and then pool

*The minimum volume that you will need for the pool is 24 $\mu$l*\    
*Always aim to pipette at least 2 $\mu$l*\    

16. Use [Illumina's online pooling calculator](https://support.illumina.com/help/pooling-calculator/pooling-calculator.htm) to combine libraries into one pool for loading onto the flow cell.\    


```{r eval = FALSE}


knitr::kable(
  target_loading,
  align = "c", 
  format = "latex",
  caption = "Targeted loading concentrations for libraries."
) %>% kableExtra::kable_styling(
  latex_options = c("HOLD")
) %>% kableExtra::column_spec(
  c(1:2),
  width = unit(75, "mm")
)
```

17. Create your multiplexed pool.\    

```{r}
target_loading <- c(rep("   ", params$libraries), "RSB with Tween 20") %>% as_tibble_col("library")

pool <- target_loading %>% mutate(
  `Vol ($\\mu$l)` = "",
)

knitr::kable(
  pool, 
  align = "c", 
  caption = "Pooling libraries", 
  format = "latex", 
  escape = F
) %>% kableExtra::kable_styling(
  latex_options = c("HOLD")
) %>% kableExtra::column_spec(
  c(1:2),
  width = unit(35, "mm")
)
```

### Prepare PhiX control

*For a ~1% PhiX control...*\    

18. In a Lo-bind tube, add 2 $\mu$l of 10 nM PhiX library to 38 $\mu$l of RSB with Tween 20.\    

19. Vortex briefly and then centrifuge at 280 g x 1'.\    

20. Keep on ice until ready to combine with pool 

### Create loading-ready pool

21. In a fresh Lo-bind tube labeled ready, combine 24 $\mu$l of pooled libraries from above with 1 $\mu$l of PhiX at 0.5 nM.\    

22. Keep on ice until ready to load the flow cell.\    

### Loading the flow cell

23. Remove cartridge *and* flow cell from 4°C and allow to sit at room temperature for at least 15' and no more than 60'.\    

24. Right before loading, remove both cartridge and flow cell from their respective foil bags.\    

25. Invert the *cartridge* ten times.

26. Remove the flow cell from its foil package (use the gray tab with the label on the tab facing up) and push to insert into the front of the cartridge. *You will hear a click*\    

27. Once inserted, pull back and remove the gray tab.\    

28. Using a new P1000 tip, pierce the Library reservoir and push the foil to the edges to enlarge the hole.\    

29. Discard the pipette tip to prevent contamination.\    

30. Add 20 $\mu$l of pool to the *bottom* of the reservoir by slowly lowering the pipette tip to the bottom of the reservoir before dispensing. *Avoid touching the foil*

### Cartridge into the machine

31. Follow instructions given by the machine.
